---
title: "Looping, Doing it Right"
author: "Brandon"
date: "12/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#https://bookdown.org/content/d1e53ac9-28ce-472f-bc2c-f499f18264a3/loops.html
rm(list=ls())
```


# What is a loop? 
Doing something over and over again

1.  FOR

   for (value in that) {
     this
    }

2.  WHILE

    while (condition) {
      code
    }
    
    
```{r}
for(i in 1:5){
  print(i) 
}

j <- 0
while(j<5){
  print(j)
  j <- j +1 #not entering information for j could results in an infinte loop
}

# I like to use FOR more than While

```

2. Using list in loops
```{r}
list1 <- c("Hello","How","Do","You","Do","Today")

for(i in list1){
  print(i)
  
}
```

3. Interrupting Loops and Skipping Elements 
```{r}
# Break
for(i in 1:5){
  if(i ==2){break}
  print(i)
}

for(i in list1){
  if(i =="Do"){break}
  print(i)
}
```


```{r}
#Next
for(i in 1:5){
  if(i ==2){next}
  print(i)
}

for(i in list1){
  if(i =="Hello"){next}
  print(i)
}

```



3. Efficient Looping? 
-   Don't make objects you dont use
-   Garbage collections
-   Initialize objects before loops
-   Use simple data-types
-   Declare Size of output prior to loops
```{r}
#   Don't make objects you dont use

for(i in 1:5){
  P1 <- i
  P2 <- 25
  P3 <- P1 + P2
  print(P3) 
  #P1 and P2 might make the code easier to understand but do no need to exsist in this loops
  
}

#do this instead
for(i in 1:5){
  P3 <-i + 25
  print(P3) 
  #P1 and P2 
  
}

```

```{r}
# Garbage collections


for(i in 1:5){
  P1 <- i
  P2 <- 25
  P3 <- P1 + P2
  print(P3) 
  # If P1 and P2 is really needed, you can remove them after the looping calculation to save memory
  # might slow down the overall calculation
  rm(P1)
  rm(P2)
  rm(P3)
}

```

```{r}
# -   Initialize objects before loops

P2 <- 25 # Declare the object before the loop to prevent multiple declarations

for(i in 1:5){
  P1 <- i
  P3 <- P1 + P2
  print(P3) 
  rm(P1)
  rm(P3)
}


```

```{r}
# -   Use simple data-types
cycle <- 1:10000

# Using a dataframe
P2 <- data.frame(25) 
system.time(
  for(i in cycle){
    P1 <- i
    P3 <- P1 + P2
  }
)

# Using a dataframe
P2 <- data.frame(25) # Remove P1 declaration
system.time(
  for(i in cycle){
    P3 <- i + P2
  }
)

#Use Numerical 
P2 <- 25 # Declare the object before the loop to prevent multiple declarations
system.time(
  for(i in cycle){
    P3 <- i + P2
  }
)

```
```{r}
# Declare Size of output prior to loops
cycle <- 1:1000000
P2 <- 25 # Declare the object before the loop to prevent multiple declarations
rm(P3)

P3 <- list()
system.time(
  for(i in cycle){
    P3[i] <- i/ P2
  }
)


P3 <- seq(1,length(cycle))
system.time(
  for(i in cycle){
    P3[i] <- i/ P2
  }
)

```


# In Summary, do as little as possible inside the loop

-  Calculate results before the loop
-  initialize objects before the loop
-  Iterate on as few numbers as possible
-  Write as little functions inside a loop as possible



4. Parallel Computing
-  Change the cycle number to understand the effects of multiple computing


```{r}
cycle <- 1:10000

system.time(
  for(i in cycle)( 
    P1 <- sqrt(i))
)
```
```{r}
library(doSNOW)
cl <- makeCluster(2, type="SOCK") # 4 – number of cores

# Using your custom defined function: "myCustomFunc()" and store in 'output' variable # Example 4
myCustomFunc <- function(i){sqrt(i)}

system.time(
  foreach(i = cycle, .combine = "c") %dopar% {
    myCustomFunc(i)
  }
)

stopCluster(cl)

```











